---
title: "Data requirements and standardization"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data requirements and standardization}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 80
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE,
  echo = TRUE,
  include = TRUE,
  fig.show = "asis",
  fig.keep = "high",
  warning = FALSE,
  message = FALSE
)
```

## Introduction

The complete workflow is illustrated as follows. This article focuses on the
package's data requirements and the first three steps, including functions
`Mapping()`, `DataCheck()` and `DataStandard()`.

``` text
data -> Mapping() -> DataCheck() -> DataStandard()
     -> prediction / diagnostic / analysis functions
```

```{r}
library(PDRobust)
data("ImperfectConSample", package = "PDRobust")
data("BiSample", package = "PDRobust")
```

Two built-in datasets are included with the package. The datasets are stored in
the `data/` directory, and the scripts used to generate them are provided in the
`data-raw/` directory.

The first dataset, `BiSample`, is a standardized longitudinal dataset that
satisfies the package’s data requirements. It contains no nonstructural missing
values; outcome values are missing only when they are structurally unobservable
due to any kind of truncation. Each row represents one subject at a specific
time. The dataset includes a subject identifier (`id`), assessment time
(`time`), survival status (`S`), treatment assignment (`A`), a binary outcome
(`Y`), and six subject-level covariates (`X1`–`X6`) .

```{r}
data("BiSample", package = "PDRobust")
head(BiSample)
```

The second one, `ImperfectConSample`, is designed to resemble longitudinal data
collected in a clinical trial with repeated follow-up assessments. Each row
represents a patient observation at a scheduled visit. The dataset includes a
patient identifier (`patient_id`), visit time (`visit_month`), survival status
(`alive_status`), treatment assignment (`treatment`), a continuous clinical
outcome, and six patient-level covariates (`X1`–`X6`) representing relevant
clinical and demographic information.

```{r}
head(ImperfectConSample)
```

## Mapping raw columns and times

`Mapping()` is the sole source of truth for structural columns, analysis window
on the raw time scale, the covariated used to estimate treatment effects, the
variables of interest, and the outcome type.

The arguments `id, time, treatment, survival, outcome` specify the corresponding
column names in the input dataset.

The arguments `baseline_time` and `cutoff_time` define the beginning and end of
the analysis window, respectively. Each must be specified as a single finite
numeric value on the raw time scale, with `baseline_time < cutoff_time`. For
example, when obeserved time points are (0, 1, ..., 4) may be set to 0 or 1,
whereas the cutoff may be set to a later time point, such as 4. For
`ImperfectConSample`, the observed time points are (0, 6, 12); therefore, the
analysis window is defined using `baseline_time = 0, cutoff_time = 12`.

The argument `covariates` and `interest_vars` are character vectors containing
the column names of the relevant covariates. Variables specified in
`interest_vars` must also be included in `covariates` . The argument `y_type`
specifies the outcome type. It should be set to `B` for a binary outcome and `C`
for a continuous outcome.

```{r}
map <- Mapping(
  id = "patient_id",
  time = "visit_month",
  treatment = "treatment",
  survival = "alive_status",
  outcome = "clinical_outcome",
  baseline_time = 0,
  cutoff_time = 12,
  covariates = paste0("X", 1:6),
  interest_vars = c("X1", "X2"),
  y_type = "C" # "B"
)

```

Users can inspect the mapping details using `print()` and the `attributes()`
function may additionally be used to inspect object-level metadata, such as its
class.

```{r}
print(map)
```

```{r}
attributes(map)
```

## Validation without modification

`DataCheck()` evaluates whether the input dataset satisfies the structural and
analytical requirements of `PDRobust` without altering the data. It identifies
potential issues, reports their severity and recommended handling, and
determines whether the dataset is ready for analysis, can be standardized using
`DataStandard()`, or requires manual resolution.

Each check includes an action-oriented message. When `strict = FALSE`, which is
the default, `DataCheck()` returns the complete validation report without
modifying the input dataset. When `strict = TRUE`, the function raises an error
if one or more failed checks are marked as analysis-blocking.

```{r}
check <- DataCheck(ImperfectConSample, map, strict = FALSE)

attributes(check)
```

`DataCheck()` returns an object of class `pd_data_check`. The object contains
the following components:

| Component | Description |
|-------------------------------|------------------------------------------------|
| `valid` | Indicates whether all checks with severity `"error"` have passed. Informational messages and warnings do not by themselves make the report invalid. |
| `ready_for_analysis` | Indicates whether the dataset can be used directly in downstream prediction, diagnostic, and analysis functions. It is `TRUE` only when no failed check is marked as analysis-blocking. |
| `manual_resolution_required` | Indicates whether at least one failed check requires manual review or correction. Such issues are not automatically resolved by `DataStandard()`. |
| `can_standardize` | Indicates whether `DataStandard()` may proceed. It is `TRUE` when no failed check requires manual resolution, although subject-level deletion may still require `drop = TRUE`. |
| `checks` | A data frame containing the itemized validation results, including the status, severity, diagnostic summary, analysis implications, and recommended handling for each check. |
| `settings` | Records the settings used during validation. The current implementation stores the validated mapping object in `settings$mapping`. |
| `diagnostics` | Contains detailed supporting information, such as affected row numbers, subject identifiers, missingness summaries, treatment-group counts, and problematic covariates. |

```{r}
check$valid
check$ready_for_analysis
check$manual_resolution_required
check$can_standardize
```

The following are diagnostics of check for `ImperfectConSample`. Users can find
detailed descriptions of all validation items in the article
*Details-for-DataCheck*.

```{r}
head(check$diagnostics)
```

## Standardization and audit attributes

`DataStandard()` returns a prepared data for later analysis.

```{r}
pd_data <- DataStandard(ImperfectConSample, map, drop = TRUE)
class(pd_data)
head(pd_data)
```

The returned `pd_data` object contains several attributes that document its
structure and the transformations applied during standardization. These
attributes can be inspected using:

```{r}
names(attributes(pd_data))
```

| Attribute | Description |
|:-------------------|:-----------------------------------------------------------|
| `names` | Stores the column names of the standardized data frame. |
| `row.names` | Stores the row identifiers used by the data frame. |
| `class` | Identifies the object classes, including its data-frame and package-specific classes. |
| `pd_mapping` | Stores the standardized mapping used by subsequent package functions, including the standardized column roles and analysis-time specification. |
| `pd_original_mapping` | Preserves the original user-supplied mapping on the raw data scale before identifiers, time points, and encodings were standardized. |
| `pd_check` | Stores the validation results associated with the standardized dataset, including readiness indicators, detected issues, and recommended handling. |
| `pd_standardization` | Records the principal transformations performed during standardization, such as identifier and time mappings, encoding conversions, retained observations, and subject-level exclusions. |

For exmaple, the original and standardized mappings can be compared using:

```{r}
attr(pd_data, "pd_original_mapping")
attr(pd_data, "pd_mapping")
```

Among these attributes, `pd_standardization` is particularly important because
it provides the primary audit trail for the changes made to the input dataset.
It can be inspected directly using:

```{r}
standardization <- attr(pd_data, "pd_standardization")
names(standardization)
```

It retains all observed assessment times within the analysis window defined by
the mapping object and transforms the ordered time grid to consecutive integers
(0, 1, ..., n).

```{r}
standardization$time_map
```

Subject identifiers are similarly mapped to consecutive integers, explicitly
recognized binary encodings are converted safely, and the resulting longitudinal
dataset is sorted by subject and standardized analysis time.

For a subject to be retained, the dataset must contain one usable record at each
retained assessment time.

```{r}
head(standardization$id_map)
```

When `drop = TRUE`, exclusions are applied at the subject level rather than to
individual records, and all exclusions are documented in the attrition report.
Outcome values that are structurally unobservable after death or other trunction
are distinguished from ordinary missing outcomes among surviving subjects and
are therefore handled separately during validation and standardization.

```{r}
standardization$attrition
```

Finally, even when a built-in dataset such as `BiSample`, or a user-supplied
dataset, already satisfies all `PDRobust` data requirements, it should still be
processed through the package’s data-preparation workflow before analysis. This
ensures that the dataset is formally validated, standardized, and supplied with
the mapping and audit attributes required by downstream functions.
