---
title: "Detailed Function Presentation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Detailed Function Presentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# PDRobust

PDRobust implements principal-stratification analyses for longitudinal outcomes
that may be truncated by death. Version 0.3.7 uses an explicit mapping-driven
workflow and refits each nuisance prediction model from the data supplied to the
current function call.

``` text
data("BiSample") -> Mapping() -> DataCheck() -> DataStandard() -> prediction / diagnostic / analysis functions
```

## 1. Load the built-in package data

```{r}
library(PDRobust)
data("BiSample", package = "PDRobust")
head(BiSample)
```

```{r}
data("ImperfectConSample", package = "PDRobust")
head(ImperfectConSample)
```

this dataset has unordered patient id, visit month是记录的时间节点；X1 –
X6是一些基础数据，是covariates；alive status记录了此刻是否存活，

Both objects are ordinary long-format data frames loaded from the package
`data/` directory. `BiSample` is the analysis-ready binary example used for the
main workflow. `ImperfectConSample` is a continuous example with deliberately
imperfect records for demonstrating validation and explicit subject-level
deletion.

## 2. Define roles and analysis settings with `Mapping()`

```{r}
mapping <- Mapping(
  id = "id",
  time = "time",
  treatment = "A",
  survival = "S",
  outcome = "Y",
  baseline_time = 0,
  cutoff_time = 2,
  covariates = c("X1", "X3", "X4", "X5", "X6"),
  interest_vars = c("X1", "X3"),
  y_type = "B"
)

mapping
```

`Mapping()` returns a `pd_mapping` object containing the structural column
names, raw baseline and cutoff times, all nuisance-model covariates, effect
modifiers, and the outcome type.

## 3. Validate the raw data with `DataCheck()`

```{r}
check <- DataCheck(BiSample, mapping, strict = FALSE)
names(check)                   
```

```{r}
check$ready_for_analysis
```

```{r}
check$manual_resolution_required
```

```{r}
head(check$checks)                   # one row per validation rule
```

```{r}
head(check$diagnostics)               # affected rows, IDs, and time summaries
```

`DataCheck()` never modifies the input data. It returns an itemized validation
report with pass/fail status, severity, analysis-blocking status, diagnostic
details, and recommended handling.

## 4. Standardize the panel with `DataStandard()`

`DataStandard()` returns a sorted `pd_data` data frame. It safely converts
explicit binary encodings, maps IDs to consecutive integers, maps the observed
analysis-time grid to consecutive integers, and attaches mapping and audit
attributes. Use `drop = TRUE` only when the reported subject-level exclusions
are intended.

```{r}
pd_data <- DataStandard(BiSample, mapping,drop =TRUE)
head(pd_data)
```

for imperfect dataset, we have:

```{r}
data("ImperfectConSample", package = "PDRobust")
head(ImperfectConSample)
```

```{r}
con_mapping <- Mapping(
  id = "patient_id",
  time = "visit_month",
  treatment = "treatment",
  survival = "alive_status",
  outcome = "clinical_outcome",
  baseline_time = 0,
  cutoff_time = 12,
  covariates = c("X1", "X2", "X3", "X4", "X5", "X6"),
  interest_vars = c("X1", "X2"),
  y_type = "C"
)

con_data <- DataStandard(ImperfectConSample, con_mapping, drop = TRUE)
```

```{r}
head(con_data)
```

```{r}
print(dim(ImperfectConSample))
print(dim(con_data))
```

## Model formulas used below

```{r}
ps_fo <- A ~ X1 + X3 + X4 + X5 + X6
prin_fo <- S ~ (X1 + X3 + X4 + X5 + X6 ) * A
out_fo <- Y ~ (X1 + X3 + X4 + X5 + X6) * A + S
```

## 5. Prediction functions

### `PSPred()`

```{r}
ps <- PSPred(
  ps_fo,
  fit_dat = pd_data,
  pred_dat = pd_data,
  mapping = mapping
)

length(ps)              # nrow(pd_data)
head(ps)
```

`PSPred()` fits a baseline logistic treatment model and returns one propensity
prediction for every row of `pred_dat`.

### `PrinPred()`

```{r}
p0 <- PrinPred(
  prin_fo,
  fit_dat = pd_data,
  pred_dat = pd_data,
  treatment = 0,
  mapping = mapping
)
head(p0)
```

`PrinPred()` returns cumulative principal-survival probabilities under the
requested treatment. In longitudinal data it fits on post-baseline rows whose
subjects survived the immediately preceding observed time. In a single-time
analysis it uses all complete rows and does not construct an at-risk indicator.

### `OutPred()`

```{r}
mu1 <- OutPred(
  out_fo,
  fit_dat = pd_data,
  pred_dat = pd_data,
  a = 1,
  mapping = mapping
)

head(mu1)
```

`OutPred()` returns predicted outcomes after setting treatment to `a` and
survival to one in `pred_dat`. It uses linear regression for `y_type = "C"` and
logistic regression for `y_type = "B"`.

## 6. Diagnostics

### `PSDiag()`

```{r}
ps_diagnostic <- PSDiag(pd_data, ps_fo)
```

```{r}
names(ps_diagnostic)
```

```{r}
print(ps_diagnostic)
ps_diagnostic$plot
```

`PSDiag()` returns unadjusted and IPTW-adjusted standardized mean differences,
ordinary IPTW weights, clipped propensity scores, plotting data, and a ggplot.
The propensity scores are always truncated internally by
`pmin(pmax(pi, 0.01), 0.99)` before weights are calculated.

### `PrinSDiag()`

```{r}
principal_diagnostic <- PrinSDiag(pd_data, ps_fo, prin_fo)

print(principal_diagnostic)
principal_diagnostic$plot
```

`PrinSDiag()` returns cutoff-aligned standardized balance statistics, clipped
propensity scores, cumulative principal scores under treatment 0 and 1, plotting
data, and a ggplot.

## 7. Principal-stratum profiling with `QR()`

`QR()` returns principal-score-weighted means and weighted quantiles of the
mapped interest variables at cutoff, together with the weights and mapping.

```{r}
principal_profile <- QR(
  pd_data,
  prin_fo,
  quantile_level = c(0.25, 0.50, 0.75)
)

print(principal_profile)
```

## 8. Treatment-group odds ratios with `ORCI()`

```{r}
or_control <- ORCI(
  pd_data,
  S ~ X1 + X3 + X4,
  a = 0,
  conf_level = 0.95
)

print(or_control)             
or_control$plot
```

`ORCI()` returns treatment-group-specific cutoff odds ratios and confidence
intervals, the fitted logistic model, analysis data, settings, and a forest
plot.

## Sensitivity analysis

```{r}
sensitivity <- SA(
  pd_data,
  ps_fo,
  prin_fo,
  out_fo,
  ratiovec = c(1, 2)
)

```

```{r}
print(sensitivity)
```

```{r}
sensitivity$plot
```

## 9. Time-specific HTEs with `HTESepT()`

```{r}
separate_hte <- HTESepT(
  pd_data,
  ps_fo = ps_fo,
  prin_fo = prin_fo,
  out_fo = out_fo,
  target_time = c(1, 2),
  B = 5,
  conf_level = 0.95,
  max_attempts = NULL,
  verbose = TRUE
)


```

```{r}
separate_hte$summary
separate_hte$forest_plot
```

```{r}
head(separate_hte$boot_mat)
```

`HTESepT()` returns an HTE estimate for each requested observed standardized
time and each mapped effect modifier, plus the intercept. `target_time` controls
reported outcome-analysis times only; principal scores still accumulate over the
full baseline-to-cutoff grid. Set `B > 0` for subject-level bootstrap standard
errors and Wald confidence intervals.

## 10. Pooled HTEs with `HTEAllT()`

```{r}
pooled_hte <- HTEAllT(
  pd_data,
  ps_fo = ps_fo,
  prin_fo = prin_fo,
  out_fo = out_fo,
  B = 0,
  verbose = FALSE
)

```

```{r}
pooled_hte$summary
pooled_hte$forest_plot
```

`HTEAllT()` always pools every observed standardized time from baseline through
cutoff. For a single analysis time it omits the time-effect term and records
that the time effect is not estimable.

## 11.

```{r}

```

`SA()` derives the outcome type from the standardized mapping. Continuous
outcomes use the original linear-model and closed-form sensitivity equations.
Binary outcomes use logistic outcome prediction and the same bounded-link HTE
equation as `HTESepT()`. In both cases it returns estimates under each specified
outcome-noise variance ratio, observed outcome variance by time, and one plot
per mapped interest variable.

The continuous built-in example can be prepared explicitly as follows:

```{r, eval=FALSE}
continuous_mapping <- Mapping(
  id = "patient_id",
  time = "visit_month",
  treatment = "treatment",
  survival = "alive_status",
  outcome = "clinical_outcome",
  baseline_time = 0,
  cutoff_time = 12,
  covariates = paste0("X", 1:6),
  interest_vars = c("X1", "X2"),
  y_type = "C"
)
continuous_data <- DataStandard(
  ImperfectConSample, continuous_mapping, drop = TRUE
)
continuous_sensitivity <- SA(
  continuous_data,
  treatment ~ X1 + X2 + X4,
  alive_status ~ X1 + X2 + X4 + treatment + visit_month,
  clinical_outcome ~ X1 + X2 + treatment,
  ratiovec = c(0, 0.05)
)
```
