---
title: "Getting started with PDRobust"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with PDRobust}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

PDRobust follows a mapping-driven workflow:

```text
data("BiSample") -> Mapping() -> DataCheck() -> DataStandard()
                 -> prediction / diagnostic / analysis functions
```

## Prepare the bundled binary data

```{r, eval=FALSE}
library(PDRobust)

data("BiSample", package = "PDRobust")
raw <- BiSample
map <- Mapping(
  id = "id",
  time = "time",
  treatment = "A",
  survival = "S",
  outcome = "Y",
  baseline_time = 0,
  cutoff_time = 2,
  covariates = c("X1", "X2", "X4"),
  interest_vars = c("X1", "X2"),
  y_type = "B"
)

check <- DataCheck(raw, map)
check$ready_for_analysis
check$checks

pd_data <- DataStandard(raw, map)
standardized_map <- attr(pd_data, "pd_mapping")
```

`Mapping()` returns a `pd_mapping`. `DataCheck()` returns a `pd_data_check`
report without modifying the data. `DataStandard()` returns a `pd_data` frame
with standardized mapping, ID/time maps, readiness checks, and attrition audit
stored as attributes.

## Prediction functions

```{r, eval=FALSE}
ps_fo <- A ~ X1 + X2 + X4
prin_fo <- S ~ X1 + X2 + X4 + A + time
out_fo <- Y ~ X1 + X2 + A

ps <- PSPred(ps_fo, pd_data, pd_data, standardized_map)
p0 <- PrinPred(prin_fo, pd_data, pd_data, 0, standardized_map)
p1 <- PrinPred(prin_fo, pd_data, pd_data, 1, standardized_map)
mu1 <- OutPred(
  out_fo,
  pd_data[pd_data$S == 1, , drop = FALSE],
  pd_data,
  a = 1,
  mapping = standardized_map
)

head(cbind(ps, p0, p1, mu1))
```

Each prediction function refits its model on `fit_dat` and returns a numeric
`pd_prediction` vector aligned with `pred_dat`.

## Diagnostics and profiling

```{r, eval=FALSE}
ps_diag <- PSDiag(pd_data, ps_fo)
prin_diag <- PrinSDiag(pd_data, ps_fo, prin_fo)
profile <- QR(pd_data, prin_fo, c(0.25, 0.5, 0.75))
odds_ratio <- ORCI(
  pd_data, S ~ X1 + X2 + X4, a = 0
)

ps_diag$smd_after
prin_diag$statistics
profile$quantile
odds_ratio$forestplotdat
```

Both diagnostic functions return ggplots. The propensity scores returned by
`PSDiag()` and `PrinSDiag()` are always clipped to `[0.01, 0.99]` before their
diagnostic weights or denominators are calculated.

## Heterogeneous treatment effects

```{r, eval=FALSE}
separate <- HTESepT(
  pd_data,
  ps_fo,
  prin_fo,
  out_fo,
  target_time = c(1, 2),
  B = 0,
  verbose = FALSE
)

pooled <- HTEAllT(
  pd_data,
  ps_fo,
  prin_fo,
  out_fo,
  B = 0,
  verbose = FALSE
)

separate$summary
separate$bootstrap_info
pooled$summary
pooled$analysis_times
```

`HTESepT()` reports estimates at explicitly selected observed times.
`HTEAllT()` pools the complete observed baseline-to-cutoff grid. Setting `B > 0`
requests successful subject-level bootstrap replications.

## Sensitivity analysis for both outcome types

```{r, eval=FALSE}
set.seed(20260728)
binary_sensitivity <- SA(
  pd_data,
  ps_fo,
  prin_fo,
  out_fo,
  ratiovec = c(0, 0.02)
)

data("ImperfectConSample", package = "PDRobust")
continuous_map <- Mapping(
  id = "patient_id",
  time = "visit_month",
  treatment = "treatment",
  survival = "alive_status",
  outcome = "clinical_outcome",
  baseline_time = 0,
  cutoff_time = 12,
  covariates = paste0("X", 1:6),
  interest_vars = c("X1", "X2"),
  y_type = "C"
)
continuous_data <- DataStandard(
  ImperfectConSample, continuous_map, drop = TRUE
)

continuous_sensitivity <- SA(
  continuous_data,
  treatment ~ X1 + X2 + X4,
  alive_status ~ X1 + X2 + X4 + treatment + visit_month,
  clinical_outcome ~ X1 + X2 + treatment,
  ratiovec = c(0, 0.05)
)

binary_sensitivity$data
continuous_sensitivity$data
```

`SA()` derives the outcome type from the standardized mapping. It returns tidy
and wide sensitivity estimates, observed outcome variances by time, and a named
plot list for the mapped interest variables.
